---
title: "bayesrules_practice"
format: html
editor: visual
---

```{r}
library(bayesrules)
library(tidyverse)
library(janitor)
```

# Chapter 1: The Big (Bayesian) Picture - 6/2

# Chapter 2 Bayes’ Rule - 6/2

```{r}
data(fake_news)

fake_news %>% 
  tabyl(type) %>% 
  adorn_totals("row")

# Tabulate exclamation usage and article type
fake_news %>% 
  tabyl(title_has_excl, type) %>% 
  adorn_totals("row")
```

### 2.1.5 Posterior Simulation

```{r}
# Define possible articles
article <- data.frame(type = c("real", "fake"))

# Define the prior model
prior <- c(0.6, 0.4)

# Simulate 3 articles
set.seed(84735)
sample_n(article, size = 3, weight = prior, replace = TRUE)

# Simulate 10000 articles. 
set.seed(84735)
article_sim <- sample_n(article, size = 10000, 
                        weight = prior, replace = TRUE)

ggplot(article_sim, aes(x = type)) + 
  geom_bar()

article_sim %>% 
  tabyl(type) %>% 
  adorn_totals("row")

article_sim <- article_sim %>% 
  mutate(data_model = case_when(type == "fake" ~ 0.2667,
                                type == "real" ~ 0.0222))

glimpse(article_sim)

# Define whether there are exclamation points
data <- c("no", "yes")

# Simulate exclamation point usage 
set.seed(3)
article_sim <- article_sim %>%
  group_by(1:n()) %>% 
  mutate(usage = sample(data, size = 1, 
                        prob = c(1 - data_model, data_model)))

article_sim %>% 
  tabyl(usage, type) %>% 
  adorn_totals(c("col","row"))

ggplot(article_sim, aes(x = type, fill = usage)) + 
  geom_bar(position = "fill")

ggplot(article_sim, aes(x = type)) + 
  geom_bar()

article_sim %>% 
  filter(usage == "yes") %>% 
  tabyl(type) %>% 
  adorn_totals("row")

ggplot(article_sim, aes(x = type)) + 
  geom_bar() + 
  facet_wrap(~ usage)
```

## 2.2 Example pop vs soda vs coke

```{r}
# Load the data
data(pop_vs_soda)

# Summarize pop use by region
pop_vs_soda %>% 
  tabyl(pop, region) %>% 
  adorn_percentages("col")
```

### 2.3.7 Posterior Simulation

```{r}
# Define possible win probabilities
chess <- data.frame(pi = c(0.2, 0.5, 0.8))

# Define the prior model
prior <- c(0.10, 0.25, 0.65)

# Simulate 10000 values of pi from the prior
set.seed(84735)
chess_sim <- sample_n(chess, size = 10000, weight = prior, replace = TRUE)

# Simulate 10000 match outcomes
chess_sim <- chess_sim %>% 
  mutate(y = rbinom(10000, size = 6, prob = pi))

# Check it out
chess_sim %>% 
  head(3)

# Summarize the prior
chess_sim %>% 
  tabyl(pi) %>% 
  adorn_totals("row")

# Plot y by pi
ggplot(chess_sim, aes(x = y)) + 
  stat_count(aes(y = ..prop..)) + 
  facet_wrap(~ pi)

# Focus on simulations with y = 1
win_one <- chess_sim %>% 
  filter(y == 1)

# Summarize the posterior approximation
win_one %>% 
  tabyl(pi) %>% 
  adorn_totals("row")

# Plot the posterior approximation
ggplot(win_one, aes(x = pi)) + 
  geom_bar()
```

# Chapter 3 The Beta Binomial Bayesian Model - 6/3

```{r}
# 3.3
summarize_beta_binomial(alpha = 45, beta = 55, y = 30, n = 50)
```

## 3.5 Simulating the Beta-Binomial

```{r}
set.seed(84735)
michelle_sim <- data.frame(pi = rbeta(10000, 45, 55)) %>% 
  mutate(y = rbinom(10000, size = 50, prob = pi))

ggplot(michelle_sim, aes(x = pi, y = y)) + 
  geom_point(aes(color = (y == 30)), size = 0.1)

# Keep only the simulated pairs that match our data
michelle_posterior <- michelle_sim %>% 
  filter(y == 30)

# Plot the remaining pi values
ggplot(michelle_posterior, aes(x = pi)) + 
  geom_density()

michelle_posterior %>% 
  summarize(mean(pi), sd(pi))

nrow(michelle_posterior)
```

# Chapter 4: Balance and Sequentiality in Bayesian Analyses

```{r}
# Import data
data(bechdel, package = "bayesrules")

# Take a sample of 20 movies
set.seed(84735)
bechdel_20 <- bechdel %>%
  sample_n(20)

bechdel_20 %>% 
  head(3)

bechdel_20 %>% 
  tabyl(binary) %>% 
  adorn_totals("row")

# Plot the Beta-Binomial model
plot_beta_binomial(alpha = 1, beta = 2, y = 4, n = 10)
plot_beta_binomial(alpha = 0.5, beta = 1, y = 4, n = 10)

# Obtain numerical summaries of the Beta-Binomial model
summarize_beta_binomial(alpha = 10, beta = 2, y = 10, n = 20)
```


# Chapter 6: Approximating the Posterior
```{r}
library(rstan)
library(bayesplot)
```

## 6.1 Grid Approximation
```{r}
# Beta Binomial Example

# Step 1: Define a grid of 6 pi values
grid_data <- data.frame(pi_grid = seq(from = 0, to = 1, length = 6))

# Step 2: Evaluate the prior & likelihood at each pi
grid_data <- grid_data %>% 
  mutate(prior = dbeta(pi_grid, 2, 2),
         likelihood = dbinom(9, 10, pi_grid))

# Step 3: Approximate the posterior
grid_data <- grid_data %>% 
  mutate(unnormalized = likelihood * prior,
         posterior = unnormalized / sum(unnormalized))

# Confirm that the posterior approximation sums to 1
grid_data %>% 
  summarize(sum(unnormalized), sum(posterior))
  sum(unnormalized) sum(posterior)
  
# Examine the grid approximated posterior
round(grid_data, 2)

# Plot the grid approximated posterior
ggplot(grid_data, aes(x = pi_grid, y = posterior)) + 
  geom_point() + 
  geom_segment(aes(x = pi_grid, xend = pi_grid, y = 0, yend = posterior))

# Set the seed
set.seed(84735)

# Step 4: sample from the discretized posterior
post_sample <- sample_n(grid_data, size = 10000, 
                        weight = posterior, replace = TRUE)

# A table of the 10000 sample values
post_sample %>% 
  tabyl(pi_grid) %>% 
  adorn_totals("row")

# Histogram of the grid simulation with posterior pdf
ggplot(post_sample, aes(x = pi_grid)) + 
  geom_histogram(aes(y = ..density..), color = "white") + 
  stat_function(fun = dbeta, args = list(11, 3)) + 
  lims(x = c(0, 1))



# Step 1: Define a grid of 101 pi values
grid_data  <- data.frame(pi_grid = seq(from = 0, to = 1, length = 101))

# Step 2: Evaluate the prior & likelihood at each pi
grid_data <- grid_data %>% 
  mutate(prior = dbeta(pi_grid, 2, 2),
         likelihood = dbinom(9, 10, pi_grid))

# Step 3: Approximate the posterior
grid_data <- grid_data %>% 
  mutate(unnormalized = likelihood * prior,
         posterior = unnormalized / sum(unnormalized))

ggplot(grid_data, aes(x = pi_grid, y = posterior)) + 
  geom_point() + 
  geom_segment(aes(x = pi_grid, xend = pi_grid, y = 0, yend = posterior))

# Set the seed
set.seed(84735)

# Step 4: sample from the discretized posterior
post_sample <- sample_n(grid_data, size = 10000, 
                        weight = posterior, replace = TRUE)

ggplot(post_sample, aes(x = pi_grid)) + 
  geom_histogram(aes(y = ..density..), color = "white", binwidth = 0.05) + 
  stat_function(fun = dbeta, args = list(11, 3)) + 
  lims(x = c(0, 1))
```

```{r}
# Gamma Poisson Example

plot_gamma_poisson(s = 3, r = 1, sum_y = 10, n = 2, posterior = FALSE)

# Step 1: Define a grid of 501 lambda values
grid_data   <- data.frame(lambda_grid = seq(from = 0, to = 15, length = 501))

# Step 2: Evaluate the prior & likelihood at each lambda
grid_data <- grid_data %>% 
  mutate(prior = dgamma(lambda_grid, 3, 1),
         likelihood = dpois(2, lambda_grid) * dpois(8, lambda_grid))

# Step 3: Approximate the posterior
grid_data <- grid_data %>% 
  mutate(unnormalized = likelihood * prior,
         posterior = unnormalized / sum(unnormalized))

# Set the seed
set.seed(84735)

# Step 4: sample from the discretized posterior
post_sample <- sample_n(grid_data, size = 10000, 
                        weight = posterior, replace = TRUE)

# Histogram of the grid simulation with posterior pdf 
ggplot(post_sample, aes(x = lambda_grid)) + 
  geom_histogram(aes(y = ..density..), color = "white") + 
  stat_function(fun = dgamma, args = list(13, 3)) + 
  lims(x = c(0, 15))
```

```{r}
# A Beta Binomial example

# STEP 1: DEFINE the model
bb_model <- "
  data {
    int<lower = 0, upper = 10> Y;
  }
  parameters {
    real<lower = 0, upper = 1> pi;
  }
  model {
    Y ~ binomial(10, pi);
    pi ~ beta(2, 2);
  }
"

# STEP 2: SIMULATE the posterior
bb_sim <- stan(model_code = bb_model, data = list(Y = 9), 
               chains = 4, iter = 5000*2, seed = 84735)

```

## 6.3 Markov chain diagnostics
```{r}
# Calculate the effective sample size ratio
neff_ratio(bb_sim, pars = c("pi"))
```


# Chapter 8: Posterior Inference and Prediction
```{r}
# Load packages
library(bayesrules)
library(tidyverse)
library(rstan)
library(bayesplot)
library(broom.mixed)
library(janitor)

# Load data
data("moma_sample")

moma_sample %>% 
  group_by(genx) %>% 
  tally()

plot_beta_binomial(alpha = 4, beta = 6, y = 14, n = 100)
```

## 8.4 Posterior Analysis with MCMC
```{r}
# STEP 1: DEFINE the model
art_model <- "
  data {
    int<lower = 0, upper = 100> Y;
  }
  parameters {
    real<lower = 0, upper = 1> pi;
  }
  model {
    Y ~ binomial(100, pi);
    pi ~ beta(4, 6);
  }
"

# STEP 2: SIMULATE the posterior
art_sim <- stan(model_code = art_model, data = list(Y = 14), 
                chains = 4, iter = 5000*2, seed = 84735)

# The actual Beta(18, 92) posterior
plot_beta(alpha = 18, beta = 92) + 
  lims(x = c(0, 0.35))
```






