---
title: "bayesian_logistic"
format: html
---

```{r}
library(tidyverse)
library(brms)
```

# 6/18 Task 3: Mess around with applying the bayesian logistic model to the data you have (a subset of it).

```{r}
# Load in 2025 season  data
mbb_pbp <-  hoopR::load_mbb_pbp(seasons = 2025)
```

## Clean 2025 Season Data
```{r}
# Clean data
# create new vars: time_elapsed and lead_diff
mbb_pbp_2025 <- mbb_pbp |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

# create new var: home_win
mbb_pbp_2025 <- mbb_pbp_2025 |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
mbb_pbp_2025 <- mbb_pbp_2025 |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)

# only need these columns
mbb_pbp_2025 <- mbb_pbp_2025 |>
  select(game_id, time_elapsed, lead_diff, home_win)

# 1 row for 1 second, no duplicates
mbb_pbp_2025_clean <- mbb_pbp_2025 |> 
  group_by(game_id) |> 
  complete(time_elapsed = 0:2400) |> 
  fill(lead_diff, home_win, .direction = "down") |> 
  ungroup() |> 
  arrange(game_id, time_elapsed, desc(row_number())) |>
  distinct(game_id, time_elapsed, .keep_all = TRUE)
```

## 2025 Data Subset
```{r}
set.seed(123)

unique_games <- mbb_pbp_2025_clean |> distinct(game_id) |> pull(game_id)

selected_games <- sample(unique_games, 50)

# subset of 50 games
subset_games <- mbb_pbp_2025_clean |> 
  filter(game_id %in% selected_games) |>
  filter(time_elapsed %% 24 == 0)

# scale time and score so algorithm runs better
subset_games <- subset_games |>
  mutate(
    time_scaled = time_elapsed / 2400,
    lead_scaled = lead_diff / 50
  )
```

## Bayesian Logistic Model
```{r}
brm_mod <- brm(home_win ~ lead_diff + time_elapsed + time_elapsed:lead_diff,
  data = subset_games,
  family = bernoulli(),
  prior = c(prior(normal(0, 2), class = "Intercept"))
)

brm_mod_scaled <- brm(
  home_win ~ lead_scaled + time_scaled + time_scaled:lead_scaled,
  data = subset_games,
  family = bernoulli(),
  prior = c(prior(normal(0, 2), class = "Intercept"))
)

plot(brm_mod)
plot(brm_mod_scaled)

summary(brm_mod)
summary(brm_mod_scaled)
```


## Compare to Frequentist Logistic Regression
```{r}
log_mod <- glm(home_win ~ time_elapsed + lead_diff + time_elapsed:lead_diff,
               data = mbb_pbp_2025_clean, family = "binomial")

tidy(log_mod)
```


## brm mod Visualization
```{r}
# model predictions
preds_brm_mod <- ggpredict(brm_mod, terms = c("time_elapsed [0:2400 by=10]", 
                             "lead_diff [-15, -10, -5, 0, 5, 10, 15]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

# visualization
ggplot(preds_brm_mod, aes(x = time_elapsed, y = predicted, 
                          color = factor(lead_diff))) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = factor(lead_diff)),
              alpha = 0.25, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff", 
       fill = "Lead Diff") +
  theme_minimal() +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 10))


# model predictions
preds_log_mod <- ggpredict(log_mod, terms = c("time_elapsed [0:2400 by=10]", 
                             "lead_diff [-15, -10, -5, 0, 5, 10, 15]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

# frequentist logistic regression visualization
ggplot(preds_log_mod, aes(x = time_elapsed, y = predicted, 
                          color = factor(lead_diff))) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = factor(lead_diff)),
              alpha = 0.25, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff", 
       fill = "Lead Diff") +
  theme_minimal() +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 10))
```


## Betting Lines/ Team Rankings Data
```{r}
library(rvest)

url_1 <- "https://www.teamrankings.com/ncaa-basketball/ranking/predictive-by-other?date=2024-11-04"
h_1 <- read_html(url_1)
tab_1 <- h_1 |> html_nodes("table")
objs_1 <- tab_1 |> html_table()
rankings_11_04 <- objs_1[[1]]

# Get all the rankings data
# 1. Build the full sequence of dates spanning your season
all_dates <- seq(as.Date("2024-11-04"), as.Date("2025-04-07"), by = "day")

# 2. Write a function that scrapes one date and returns a tibble (or NULL on failure)
scrape_one_date <- function(date) {
  url <- paste0("https://www.teamrankings.com/ncaa-basketball/ranking/predictive-by-other?date=", date)
  
  result <- tryCatch({
    h <- read_html(url)
    tab <- h |> html_nodes("table")
    objs <- tab |> html_table()
    rankings <- objs[[1]] |> mutate(rank_date = date)
    rankings
  }, error = function(e) {
    message("Failed on date: ", date, " — ", e$message)
    NULL
  })
  
  Sys.sleep(1)  # be polite to the server, avoid hammering it with rapid requests
  return(result)
}

# 3. Loop over every date, scraping one at a time
rankings_list <- map(all_dates, scrape_one_date)

# 4. Drop any failed dates (NULLs) and combine everything into one tibble
all_rankings <- rankings_list |> 
  compact() |>     # removes NULL entries from failed scrapes
  bind_rows()


# clean up team name column
all_rankings <- all_rankings |> select(Rank, Team, Rating, rank_date) |>
  mutate(Team = str_remove(Team, "\\s*\\(\\d+-\\d+\\)$"))
```

## Fix the mismatched names
```{r}
# determine if name or alt name is better to join with
main_teams <- mbb_pbp |> distinct(home_team_name) |> pull(home_team_name)
main_teams_alt <- mbb_pbp |> distinct(home_team_name_alt) |> pull(home_team_name_alt)
rank_teams <- all_rankings |> distinct(Team) |> pull(Team)

# teams in your main data with no match in rankings
setdiff(main_teams, rank_teams)
setdiff(main_teams_alt, rank_teams)


library(fuzzyjoin)

missing_teams <- setdiff(main_teams_alt, rank_teams)

missing_df <- tibble(team_main = missing_teams)
rank_df <- tibble(team_rank = rank_teams)

# look through the mismatched names
fuzzy_matches <- stringdist_join(
  missing_df, rank_df,
  by = c(team_main = "team_rank"),
  method = "jw",        # Jaro-Winkler, works well for name variants
  max_dist = 0.15,       
  distance_col = "dist"
) |>
  arrange(team_main, dist)

fuzzy_matches |> group_by(team_main) |> slice_min(dist, n = 1) |> print(n = 100)

# Manually fix the mismatched names
name_fixes <- tribble(
  ~team_main,           ~team_rank,
  "App St",         "App State",
  "Charleston Southern",    "Charleston So",
  "Coastal Carolina",    "Coastal Car",
  "East Tennessee St",    "E Tennessee St",
  "Florida International",    "Florida Intl",
  "Georgia Southern",    "Georgia So",
  "Loyola Chicago",    "Loyola Chi",
  "Loyola Marymount",    "Loyola Mymt",
  "Miami (OH)",    "Miami OH",
  "Middle Tennessee",    "Middle Tenn",
  "Mississippi Valley St",    "Miss Valley St",
  "Mount St. Mary's",    "Mt St Mary's",
  "NC St",    "NC State",
  "North Alabama",    "N Alabama",
  "North Carolina A&T",    "NC A&T",
  "North Carolina Central",    "NC Central",
  "North Texas",    "N Texas",
  "Prairie View A&M",    "Prairie View",
  "SIU Edwardsville",    "SIU Edward",
  "San José St",    "San Jose St",
  "Seattle U",    "Seattle",
  "South Alabama",    "S Alabama",
  "South Carolina St",    "S Carolina St",
  "St. Bonaventure",    "St Bonaventure",
  "St. John's",    "St John's",
  "Texas Southern",    "Texas So",
  "UAlbany",    "Albany",
  "UNC Asheville",    "NC Asheville",
  "UNC Greensboro",    "NC Greensboro",
  "UNC Wilmington",    "NC Wilmington",
  "UT Rio Grande Valley",    "UT Rio Grande",
  "West Georgia",    "W Georgia"
  
)

mbb_pbp |>
  left_join(name_fixes, by = c("home_team_name_alt" = "team_main")) |>
  mutate(home_team_join = coalesce(team_rank, home_team_name_alt)) |>
  select(-team_rank)
```

## Joining team ranking variables with mbb_pbp
```{r}
# add new home and away team names
mbb_pbp_rank <- mbb_pbp |>
  left_join(name_fixes, by = c("home_team_name_alt" = "team_main")) |>
  mutate(home_team_join = coalesce(team_rank, home_team_name_alt)) |>
  select(-team_rank) |>
  left_join(name_fixes, by = c("away_team_name_alt" = "team_main")) |>
  mutate(away_team_join = coalesce(team_rank, away_team_name_alt)) |>
  select(-team_rank)


library(data.table)

main_dt <- as.data.table(mbb_pbp_rank)
rank_dt <- as.data.table(all_rankings)  # has Team, Rating, Rank, rank_date

setkey(rank_dt, Team, rank_date)

# home team ratings
setkey(main_dt, home_team_join, game_date)  # adjust game_date to your actual date column
main_dt <- rank_dt[main_dt, on = .(Team = home_team_join, rank_date = game_date), roll = TRUE]
setnames(main_dt, c("Rating", "Rank"), c("home_rating", "home_rank"))
main_dt[, Team := NULL]  # drop the join key column, no longer needed

# away team ratings
setkey(main_dt, away_team_join, rank_date)  # rank_date currently holds game_date here
main_dt <- rank_dt[main_dt, on = .(Team = away_team_join, rank_date = rank_date), roll = TRUE]
setnames(main_dt, c("Rating", "Rank"), c("away_rating", "away_rank"))
main_dt[, Team := NULL]

# add the time, lead, home win columns first, and game_date got removed
final_data <- main_dt |>
  as_tibble() |>
  mutate(rating_diff = home_rating - away_rating) |>
  select(game_id, game_date, time_elapsed, lead_diff, home_win, rating_diff)
```


# 6/23 Bayesian Logistic Model with Ratings Data

## Attempt 2: Joining pbp with rankings data
```{r}
library(tidyverse)
library(fuzzyjoin)
library(data.table)

## ---- STEP 1: your existing play-by-play cleaning ----
mbb_pbp_2025 <- mbb_pbp |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

mbb_pbp_2025 <- mbb_pbp_2025 |> 
  group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup()

mbb_pbp_2025 <- mbb_pbp_2025 |> 
  mutate(
    time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
    end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
  )

## keep the columns you need for the ranking join too — adjust names as needed
mbb_pbp_2025 <- mbb_pbp_2025 |>
  select(game_id, game_date, home_team_name_alt, away_team_name_alt,
         time_elapsed, lead_diff, home_win)

## ---- STEP 2: clean rankings data ----
all_rankings <- all_rankings |> 
  select(Rank, Team, Rating, rank_date) |>
  mutate(Team = str_remove(Team, "\\s*\\(\\d+-\\d+\\)$"))

## ---- STEP 3: fix mismatched team names ----
name_fixes <- tribble(
  ~team_main,                  ~team_rank,
  "App St",                    "App State",
  "Charleston Southern",       "Charleston So",
  "Coastal Carolina",          "Coastal Car",
  "East Tennessee St",         "E Tennessee St",
  "Florida International",     "Florida Intl",
  "Georgia Southern",          "Georgia So",
  "Loyola Chicago",            "Loyola Chi",
  "Loyola Marymount",          "Loyola Mymt",
  "Miami (OH)",                "Miami OH",
  "Middle Tennessee",          "Middle Tenn",
  "Mississippi Valley St",     "Miss Valley St",
  "Mount St. Mary's",          "Mt St Mary's",
  "NC St",                     "NC State",
  "North Alabama",             "N Alabama",
  "North Carolina A&T",        "NC A&T",
  "North Carolina Central",    "NC Central",
  "North Texas",               "N Texas",
  "Prairie View A&M",          "Prairie View",
  "SIU Edwardsville",          "SIU Edward",
  "San José St",               "San Jose St",
  "Seattle U",                 "Seattle",
  "South Alabama",             "S Alabama",
  "South Carolina St",         "S Carolina St",
  "St. Bonaventure",           "St Bonaventure",
  "St. John's",                "St John's",
  "Texas Southern",            "Texas So",
  "UAlbany",                   "Albany",
  "UNC Asheville",             "NC Asheville",
  "UNC Greensboro",            "NC Greensboro",
  "UNC Wilmington",            "NC Wilmington",
  "UT Rio Grande Valley",      "UT Rio Grande",
  "West Georgia",              "W Georgia",
  "Florida Gulf Coast",     "FGCU",
  "Northern Arizona",     "N Arizona",
  "Ole Miss",     "Mississippi",
  "American University",     "American",
  "UC San Diego",     "UCSD",
  "George Washington",     "G Washington",
  "Bethune-Cookman",     "Bethune",
  "Abilene Christian",     "Abl Christian",
  "Cal St",     "CS Northridge",
  "East Carolina",     "E Carolina",
  "California Baptist",     "Cal Baptist",
  "Massachusetts",     "UMass",
  "UC Santa Barbara",     "UCSB",
  "Northern Colorado",     "N Colorado",
  "Southern Utah",     "S Utah",
  "Central Connecticut",     "C Connecticut",
  "Western Kentucky",     "W Kentucky",
  "Northwestern St",     "NW State",
  "Texas A&M-Corpus Christi",     "Texas A&M-CC",
  "James Madison",     "J Madison",
  "Long Island University",     "LIU",
  "Mt St. Mary's",     "Mt St Mary's",
  "Arkansas-Pine Bluff",     "AR-Pine Bluff",
  "Southeast Missouri St",     "SE Missouri St",
  "St. Thomas-Minnesota",     "St Thomas",
  "North Dakota St",     "N Dakota St",
  "Central Michigan",     "C Michigan",
  "South Dakota St",     "S Dakota St",
  "Northern Iowa",     "N Iowa",
  "Southern Illinois",     "S Illinois",
  "Boston University",     "Boston U",
  "UIC",     "Illinois Chicago",
  "Tennessee Tech",     "Tenn Tech",
  "Saint Francis",     "St Francis PA",
  "Fairleigh Dickinson",     "F Dickinson",
  "Loyola Maryland",     "Loyola MD",
  "Northern Kentucky",     "N Kentucky",
  "Western Michigan",     "W Michigan",
  "Northern Illinois",     "N Illinois",
  "Stephen F. Austin",     "SF Austin",
  "East Texas A&M",     "E Texas A&M",
  "South Florida",     "S Florida",
  "South Carolina Upstate",     "SC Upstate",
  "IU Indianapolis",     "IU Indy",
  "Maryland Eastern Shore",     "Maryland ES",
  "Eastern Illinois",     "E Illinois",
  "Eastern Washington",     "E Washington",
  "Eastern Michigan",     "E Michigan",
  "Purdue Fort Wayne",     "Purdue FW",
  "North Florida",     "N Florida",
  "Western Carolina",     "W Carolina",
  "Western Illinois",     "W Illinois",
  "Central Arkansas",     "C Arkansas",
  "Houston Christian",     "Hou Christian",
  "Pennsylvania",     "Penn",
  "Southern Indiana",     "S Indiana",
  "Queens University",     "Queens",
  "Eastern Kentucky",     "E Kentucky"
)

mbb_pbp_2025 <- mbb_pbp_2025 |>
  left_join(name_fixes, by = c("home_team_name_alt" = "team_main")) |>
  mutate(home_team_join = coalesce(team_rank, home_team_name_alt)) |>
  select(-team_rank) |>
  left_join(name_fixes, by = c("away_team_name_alt" = "team_main")) |>
  mutate(away_team_join = coalesce(team_rank, away_team_name_alt)) |>
  select(-team_rank)

all_rankings <- all_rankings |> 
  filter(!Team %in% c("NW St", "Mt St. Mary's"))

## ---- STEP 4 (do this before the join): handle remaining non-D1 teams ----
## After fixes, check what's still unmatched:
rank_teams <- all_rankings |> distinct(Team) |> pull(Team)
still_missing_home <- setdiff(unique(mbb_pbp_2025$home_team_join), rank_teams)
still_missing_away <- setdiff(unique(mbb_pbp_2025$away_team_join), rank_teams)
still_missing <- union(still_missing_home, still_missing_away)
still_missing  # review this list — confirm these are genuinely non-D1




## once confirmed, append a fallback rating for them into all_rankings
## so the rolling join below has something to find
min_rating <- min(all_rankings$Rating, na.rm = TRUE)

non_d1_fill <- tibble(
  Team = still_missing,
  Rating = min_rating - 1,
  Rank = NA_integer_,
  rank_date = min(all_rankings$rank_date)  # so it's always available for rolling join
)

all_rankings <- bind_rows(all_rankings, non_d1_fill)

## ---- STEP 5: rolling join on date, for both home and away ----
main_dt <- as.data.table(mbb_pbp_2025)
rank_dt <- as.data.table(all_rankings)

setkey(rank_dt, Team, rank_date)

# home team ratings
setkey(main_dt, home_team_join, game_date)
main_dt <- rank_dt[main_dt, on = .(Team = home_team_join, rank_date = game_date), roll = TRUE]
setnames(main_dt, c("Rating", "Rank"), c("home_rating", "home_rank"))
main_dt[, Team := NULL]

# away team ratings — note rank_date currently holds game_date from the prior join
setkey(main_dt, away_team_join, rank_date)
main_dt <- rank_dt[main_dt, on = .(Team = away_team_join, rank_date = rank_date), roll = TRUE]
setnames(main_dt, c("Rating", "Rank"), c("away_rating", "away_rank"))
main_dt[, Team := NULL]

## ---- STEP 6: build rating_diff and trim to final columns ----
final_data <- main_dt |>
  as_tibble() |>
  mutate(rating_diff = home_rating - away_rating) |>
  select(game_id, time_elapsed, lead_diff, home_win, rating_diff)

## sanity checks before trusting it
sum(is.na(final_data$rating_diff))   # should be 0 (or very close) if joins worked
final_data |> filter(game_id == final_data$game_id[1])  # spot-check one game by hand
# check one game at a time
View(final_data |> filter(game_id == 401746082))
View(mbb_pbp |> filter(game_id == 401727065))

# check rating for mt st marys and nw st
all_rankings |> filter(str_detect(Team, "NW|Northwestern")) |> distinct(Team)
all_rankings |> filter(str_detect(Team, "Mt St|Mount St")) |> distinct(Team)

# check if they have actual rows with ratings
View(all_rankings |> filter(Team == "Baylor"))
View(all_rankings |> filter(Team == "Mt St Mary's"))

# see what home_team_join actually says for these teams
mbb_pbp_2025 |> 
  filter(str_detect(home_team_name_alt, "NW|Northwestern|Mount|Mt St")) |>
  distinct(home_team_name_alt, home_team_join)

mbb_pbp_2025 |> 
  filter(str_detect(away_team_name_alt, "NW|Northwestern|Mount|Mt St")) |>
  distinct(away_team_name_alt, away_team_join)

```


## Bayesian logistic model with final_data
```{r}
# 1 row for 1 second, no duplicates
final_data <- final_data |> 
  group_by(game_id) |> 
  complete(time_elapsed = 0:2400) |> 
  fill(lead_diff, home_win, rating_diff, .direction = "down") |> 
  ungroup() |> 
  arrange(game_id, time_elapsed, desc(row_number())) |>
  distinct(game_id, time_elapsed, .keep_all = TRUE)

# save final data
# saveRDS(final_data, file = "final_data.rds")


# Data Subset
set.seed(123)

unique_games <- final_data |> distinct(game_id) |> pull(game_id)

selected_games <- sample(unique_games, 50)

# subset of 50 games
subset_games <- final_data |> 
  filter(game_id %in% selected_games) |>
  filter(time_elapsed %% 24 == 0)

# scale time and score so algorithm runs better
subset_games <- subset_games |>
  mutate(
    time_scaled = time_elapsed / 2400,
    lead_scaled = lead_diff / 50,
    ratings_scaled = rating_diff / 50
  )


# Model
brm_mod_rank <- brm(home_win ~ lead_diff + time_elapsed + rating_diff + time_elapsed:lead_diff + time_elapsed:rating_diff,
  data = subset_games,
  family = bernoulli(),
  prior = c(prior(normal(0, 2), class = "Intercept"))
)

plot(brm_mod)
summary(brm_mod)
round(fixef(brm_mod_rank), 4)

# Plot
# model predictions
preds_brm_mod_rank <- ggpredict(brm_mod_rank, terms = c("time_elapsed [0:2400 by=10]", 
                             "lead_diff [-15, -10, -5, 0, 5, 10, 15]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

# visualization
# start at 5 min, lead diff order was flipped before, maybe b/c not enough data
ggplot(preds_brm_mod_rank, aes(x = time_elapsed, y = predicted, 
                          color = factor(lead_diff))) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = factor(lead_diff)),
              alpha = 0.25, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff", 
       fill = "Lead Diff") +
  theme_minimal()  +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 10))
  
```


## .rds file from ada

```{r}
# this is just for 50 games, same as above
log_test <- readRDS("bayesian_logistic.rds")
```

```{r}
# 25% of the dataset, 1534 games
log_quarter <- readRDS("bayesian_logistic_quarter.rds")

ggplot(log_quarter, aes(x = time_elapsed, y = predicted, 
                          color = factor(lead_diff))) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = factor(lead_diff)),
              alpha = 0.25, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff", 
       fill = "Lead Diff") +
  theme_minimal()
```

```{r}
# 50% of the dataset, 3068 games
log_half <- readRDS("bayesian_logistic_half.rds")

ggplot(log_half, aes(x = time_elapsed, y = predicted, 
                          color = factor(lead_diff))) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = factor(lead_diff)),
              alpha = 0.25, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff", 
       fill = "Lead Diff") +
  theme_minimal()
```

```{r}
# All of the dataset, 6136 games
log_full <- readRDS("bayesian_logistic_full.rds")

ggplot(log_full, aes(x = time_elapsed, y = predicted, 
                          color = factor(lead_diff))) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = factor(lead_diff)),
              alpha = 0.25, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff", 
       fill = "Lead Diff") +
  theme_minimal()
```





