---
title: "Closeread Test"
format:
  closeread-html:
    embed-resources: true
    cr-style:
      narrative-background-color-overlay: darkslategrey
      narrative-text-color-overlay: "#e2e2e2"
---

```{r}
#| include: false
library(tidyverse)
duke_uconn_db <- read_csv("duke_uconn_db.csv")
```

## Duke vs UConn - 2026 Elite Eight

This is a test Closeread document using the Dynamic Bayes win probability estimator applied to the Duke vs UConn Elite Eight game.

:::{.cr-section}

Duke was the home team. The dashed horizontal line marks 50% win probability and the dashed vertical line marks halftime. @cr-game-plot


Halfway through the game Duke had built a comfortable lead and their win probability peaked at 98.6% when they had a 19 point lead just before the 1st half ended. [@cr-game-plot]{pan-to="0%,70%" scale-by="2.5"}

In the final seconds of the game, UConn's freshman Braylon Mullins hits a 35-foot buzzer beating three pointer to win. Just before the shot Duke's win probability was at 88%. [@cr-game-plot]{pan-to="-100%,25%" scale-by="2"}

:::{#cr-game-plot}

```{r}
#| warning: false

ggplot(duke_uconn_db, aes(x = time_elapsed, y = win_prob)) +
  geom_line(linewidth = 0.4) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_x_continuous(breaks = seq(0, 2400, by = 300)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Duke vs UConn Dynamic Bayes Win Probability",
       x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability (Duke)") +
  theme_minimal(base_size = 14)
```

:::

:::