---
title: "data_exploration"
format: html
---

```{r}
library(tidyverse)
```

# Initial Data Collection

```{r}
# You can install using the pak package using the following code:
if (!requireNamespace('pak', quietly = TRUE)){
  install.packages('pak')
}
pak::pak("sportsdataverse/hoopR")
```

```{r}
# pbp for 2026 season
tictoc::tic()
progressr::with_progress({
  mbb_pbp <-  hoopR::load_mbb_pbp(seasons = 2026)
})
tictoc::toc()

# for all seasons: load_mbb_pbp(seasons = 2002:2026), but file is too large

print(object.size(mbb_pbp_2026), units = "auto")
```


# Initial Data Exploration

```{r}
# only keep important cols and add time_elapsed and lead_diff 
mbb_pbp_2026 <- mbb_pbp |> select(
  game_id, home_score, away_score, end_game_seconds_remaining, scoring_play, score_value, home_team_name, away_team_name, season_type, period_number, game_date 
  ) |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)
```

```{r}
# duke vs uconn game
duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_name == "Duke" & away_team_name == "UConn")

```

```{r}
# There are 6275 different games from the 2026 season and there are some games that involve one non D1 school
mbb_pbp_2026 |> group_by(game_id)

```

```{r}
# home win rate
mbb_pbp_2026 |> 
  group_by(game_id) |>
  filter(time_elapsed == max(time_elapsed, na.rm = TRUE)) |>
  slice(1) |> 
  ungroup() |>
  summarise(home_win_rate = mean(home_score > away_score))

# for 2026: home win rate is 63.7% which is lower than expected (google says 68.7%)
# for 2025: home win rate is 63.4% which is lower than expected
# but not sure this calculation is correct because the home_win variable mean is 0.666 but that is by play rather than game.
```

```{r}
# duke vs uconn plot

duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_name == "Duke" & away_team_name == "UConn")

duke_uconn |>
  ggplot(aes(x = time_elapsed, y = lead_diff)) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal()
```

```{r}
# find games with least/most plays to see if there is scraping errors
mbb_pbp_2026 |>
  group_by(game_id) |>
  summarise(n_plays = n()) |>
  arrange(n_plays)

# Game with least plays: the issue is that espn seems to be missing plays and not an issue with the scraping
mbb_pbp_2026 |> filter(game_id == 401824072) |> relocate(home_team_name, away_team_name)

# Game with most plays: the issue is that espn play by play glitched and it doubled up every play
mbb_pbp_2026 |> filter(game_id == 401817534) |> relocate(home_team_name, away_team_name)

```

```{r}
# what happens for an overtime game
ot_game <- mbb_pbp_2026 |> 
  filter(home_team_name == "Kentucky" & away_team_name == "Santa Clara") |>
  relocate(time_elapsed) |> arrange(desc(time_elapsed))
# for time elapsed there is one NA value
# then when in ot, end_game_seconds_remaining resets to 300 (5 mins), 
# for example then there are two very different game scores when time elapsed is 2400, it is 73-73 and 89-84

ot_game |>
  ggplot(aes(x = time_elapsed, y = lead_diff)) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal()
```


# Frequentist Logistic Regression Model

```{r}
# Get data for 2025 season
tictoc::tic()
progressr::with_progress({
  mbb_pbp_2025 <-  hoopR::load_mbb_pbp(seasons = 2025)
})
tictoc::toc()
```

```{r}
# create new vars: time_elapsed & lead_diff
mbb_pbp_2025 <- mbb_pbp_2025 |> select(
  game_id, home_score, away_score, end_game_seconds_remaining, scoring_play, score_value, home_team_name, away_team_name, season_type, period_number, game_date 
  ) |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

# create new var: home_win
mbb_pbp_2025 <- mbb_pbp_2025 |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
mbb_pbp_2025 <- mbb_pbp_2025 |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)
```

```{r}
library(broom)
library(ggeffects)
```

```{r}
# logistic regression model
log_mod <- glm(home_win ~ time_elapsed + lead_diff + time_elapsed:lead_diff,
               data = mbb_pbp_2025, family = "binomial")

log_mod |> tidy()
```

```{r}
# model predictions
preds_log_mod <- ggpredict(log_mod, terms = c("time_elapsed [0:2400 by=10]", 
                             "lead_diff [-15, -10, -5, 0, 5, 10, 15]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

# visualization
ggplot(data = preds_log_mod, aes(x = time_elapsed, y = predicted)) +
  geom_line(aes(color = factor(lead_diff)), linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff") +
  theme_minimal()
```
Plot Takeaways
- When lead_diff = 0, it starts at 67% and slowly decreases over time, so home court advantage slightly fades as time goes on
- The lines spread out over time because of the interaction term, for example a 5 point lead means more in the last minutes of the game then the first minutes.
- Home court advantage shifts all of the lines up, the positive lead diff lines are higher up than the negative lead diff lines are lower down.


```{r}
# heatmap visualization

preds_heatmap <- ggpredict(log_mod, terms = c("time_elapsed [0:2400 by=10]", 
                                               "lead_diff [-25:25 by=1]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

ggplot(preds_heatmap, aes(x = time_elapsed, y = lead_diff, fill = predicted)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5) +
  labs(x = "Time Elapsed (seconds)", 
       y = "Home Team Lead Diff",
       fill = "Home Win Prob") +
  theme_minimal()
```

```{r}
# duke vs uconn visualization

duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_name == "Duke" & away_team_name == "UConn") |>
  mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
) |>
  mutate(
    predicted = predict(log_mod, newdata = pick(everything()), type = "response"
                        ))

ggplot(duke_uconn, aes(x = time_elapsed, y = predicted)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       title = "Duke vs UConn") +
  theme_minimal()

```
- The peak home team win prob was 98% right before halftime which is very similar to ESPN win probability model.



# Basic Data Summary

```{r}
mbb_pbp |> summarise(
  total_games = n_distinct(game_id),
  total_playoff_games = n_distinct(game_id[season_type == 3]),
  total_teams = n_distinct(home_team_id)
  )
```
- total games: 6275
  - So with a row every second there will be over 15 million rows, maybe just pull teams from the biggest conferences.
- total playoff games: 105 (includes other tournaments besides MM)
- total teams: using home_team_id = 366, using away_team_id = 720
  - There are 365 D1 basketball teams, so some non D1 schools are included in the data.


# Clean Duplicate and Missing Rows for time_elapsed
```{r}
# first test on duke_uconn game, get duke_uconn game data
mbb_pbp <-  hoopR::load_mbb_pbp()

mbb_pbp_2026 <- mbb_pbp |> select(
  game_id, home_score, away_score, end_game_seconds_remaining, scoring_play, score_value, home_team_name, away_team_name, season_type, period_number, game_date 
  ) |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_name == "Duke" & away_team_name == "UConn") |>
  mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)
```

- before cleaning: only 463 rows and there are duplicate rows

```{r}
# clean data, 1 row for 1 second, no duplicates
duke_uconn_clean <- duke_uconn |>
  select(c(12,13,2,3)) |>
  complete(time_elapsed = 0:2400) |>
  fill(home_score, away_score, lead_diff, .direction = "down") |>
  group_by(time_elapsed) |>
  slice_tail(n = 1) |>
  ungroup()
```


# Bracket Testing
```{r}
# 1. Define team slots and coordinates for a round
# (You would scale this up for all 64 teams, or just map the final 16/8 to keep it clean)
bracket_data <- data.frame(
  team = c("Duke", "UConn", "Michigan", "Virginia"),
  x_start = c(1, 1, 5, 5),   # Left and right sides of the screen
  y_start = c(4, 3, 4, 3),
  round_reached = c(3, 4, 4, 1) # 1=R64, 2=R32, 3=S16, 4=E8, etc.
)

# 2. To create your 6 separate versions, you just filter the data!
# Version 1: No games played (Show everyone at round 1)
bracket_v1 <- bracket_data

# Version 2: Sweet 16 played (Only draw paths for teams who reached round 3 or higher)
bracket_v4 <- bracket_data %>% 
  mutate(team_label = ifelse(round_reached >= 3, team, ""))

# 3. Plot using standard ggplot lines
ggplot(bracket_v4) +
  geom_segment(aes(x = x_start, y = y_start, xend = x_start + 1, yend = y_start)) +
  geom_text(aes(x = x_start + 0.1, y = y_start + 0.1, label = team_label)) +
  theme_void() # Removes background grids for a clean look


# 1. Generate the initial bracket plot
p <- plotTourn(games = 64)

# 2. Filter out all layers containing text geom
p$layers <- purrr::keep(p$layers, ~ !inherits(.x$geom, "GeomText"))

# 3. View the clean bracket
print(p)
```

# First Round Bracket
```{r}
# Could try an r package: "brackteer", "mRchmadness", or "MMBracketR"

# MMBracketR
#install.packages("remotes")
#remotes::install_github("alexkaechele/MMBracketR")

library(MMBracketR)

plotTourn()



# 1. Start with the base package layout
bracket <- plotTourn(games = 64) +
  # 2. Add professional styling to match your win probability plots
  theme_void() + 
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16, margin = margin(b = 10))
  ) +
  labs(title = "2026 NCAA Tournament Bracket")

bracket$layers <- purrr::keep(bracket$layers, ~ !inherits(.x$geom, "GeomText"))

# 3. Use annotate() to layer team names onto the canvas segments
# (You can tweak the X/Y coordinates to place them exactly on the lines)
bracket_first_round <- bracket +
  # East Region
  annotate("text", x = 0.5, y = 32.3, 
           label = "Duke (1)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 31.3, 
           label = "Sienna (16)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 30.3, 
           label = "Ohio St. (8)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 29.3, 
           label = "TCU (9)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 28.3, 
           label = "St. John's (5)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 27.3, 
           label = "Northern Iowa (12)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 26.3, 
           label = "Kansas (4)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 25.3, 
           label = "Cal Baptist (13)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 24.3, 
           label = "Louisville (6)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 23.3, 
           label = "South Florida (11)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 22.3, 
           label = "Michigan St. (3)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 21.3, 
           label = "NDSU (14)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 20.3, 
           label = "UCLA (7)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 19.3, 
           label = "UCF (10)", size = 1.5, color = "black") +
  annotate("label", x = 0.5, y = 18.3, 
           label = "UConn (2)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  annotate("text", x = 0.5, y = 17.3, 
           label = "Furman (15)", size = 1.5, 
           fontface = "bold", color = "black") +
  
  # South Region
  annotate("text", x = 0.5, y = 16.3, 
           label = "Flordia (1)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 15.3, 
           label = "Prairie View A&M (16)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 14.3, 
           label = "Clemson (8)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 13.3, 
           label = "Iowa (9)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 12.3, 
           label = "Vanderbilt (5)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 11.3, 
           label = "McNeese (12)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 10.3, 
           label = "Nebraska (4)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 9.3, 
           label = "Troy (13)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 8.3, 
           label = "UNC (6)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 7.3, 
           label = "VCU (11)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 6.3, 
           label = "Illinois (3)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 5.3, 
           label = "Penn (14)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 4.3, 
           label = "Saint Mary's (7)", size = 1.5, color = "black") +
  annotate("text", x = 0.5, y = 3.3, 
           label = "Texas A&M (10)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 2.3, 
           label = "Houston (2)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 0.5, y = 1.3, 
           label = "Idaho (15)", size = 1.5, 
           fontface = "bold", color = "black") +
  
  # West Region
  annotate("text", x = 9.5, y = 32.3, 
           label = "Arizona (1)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 31.3, 
           label = "Long Island (16)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 30.3, 
           label = "Villanova (8)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 29.3, 
           label = "Utah St. (9)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 28.3, 
           label = "Wisconsin (5)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 27.3, 
           label = "High Point (2)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 26.3, 
           label = "Arkansas (4)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 25.3, 
           label = "Hawaii (13)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 24.3, 
           label = "BYU (6)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 23.3, 
           label = "Texas (11)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 22.3, 
           label = "Gonazaga (3)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 21.3, 
           label = "Kennesaw St. (14)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 20.3, 
           label = "Miami FL (7)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 19.3, 
           label = "Missouri (10)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 18.3, 
           label = "Purdue (2)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 17.3, 
           label = "Queens (15)", size = 1.5, color = "black") +
  
  # Midwest Region
  annotate("label", x = 9.5, y = 16.3, 
           label = "Michigan (1)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  annotate("text", x = 9.5, y = 15.3, 
           label = "Howard (16)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 14.3, 
           label = "Georgia (8)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 13.3, 
           label = "Saint Louis (9)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 12.3, 
           label = "Texas Tech (5)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 11.3, 
           label = "Akron (12)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 10.3, 
           label = "Alabama (4)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 9.3, 
           label = "Hofstra (13)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 8.3, 
           label = "Tennessee (6)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 7.3, 
           label = "Miami OH (11)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 6.3, 
           label = "Virginia (3)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 5.3, 
           label = "Wright St. (14)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 4.3, 
           label = "Kentucky (7)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 3.3, 
           label = "Santa Clara (10)", size = 1.5, color = "black") +
  annotate("text", x = 9.5, y = 2.3, 
           label = "Iowa St. (2)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 9.5, y = 1.3, 
           label = "Tennessee St. (15)", size = 1.5, color = "black")
  

bracket_first_round
```

# Second Round Bracket
```{r}
bracket_second_round <- bracket_first_round +
  # East Region
  annotate("text", x = 1.5, y = 31.8, 
           label = "Duke (1)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 1.5, y = 29.8, 
           label = "TCU (9)", size = 1.5, color = "black") +
  annotate("text", x = 1.5, y = 27.8, 
           label = "St. John's (5)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 1.5, y = 25.8, 
           label = "Kansas (4)", size = 1.5, color = "black") +
  annotate("text", x = 1.5, y = 23.8, 
           label = "Louisville (6)", size = 1.5, color = "black") +
  annotate("text", x = 1.5, y = 21.8, 
           label = "Michigan St. (3)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 1.5, y = 19.8, 
           label = "UCLA (7)", size = 1.5, color = "black") +
  annotate("label", x = 1.5, y = 17.8, 
           label = "UConn (2)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  
  # South Region
  annotate("text", x = 1.5, y = 15.8, 
           label = "Florida (1)", size = 1.5, color = "black") +
  annotate("text", x = 1.5, y = 13.8, 
           label = "Iowa (9)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 1.5, y = 11.8, 
           label = "Vanderbilt (5)", size = 1.5, color = "black") +
  annotate("text", x = 1.5, y = 9.8, 
           label = "Nebraska (4)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 1.5, y = 7.8, 
           label = "VCU (11)", size = 1.5, color = "black") +
  annotate("text", x = 1.5, y = 5.8, 
           label = "Illinois (2)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 1.5, y = 3.8, 
           label = "Texas A&M (10)", size = 1.5, color = "black") +
  annotate("text", x = 1.5, y = 1.8, 
           label = "Houston (2)", size = 1.5, 
           fontface = "bold", color = "black") +
  
  # West Region
  annotate("text", x = 8.5, y = 31.8, 
           label = "Arizona (1)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 8.5, y = 29.8, 
           label = "Utah St. (9)", size = 1.5, color = "black") +
  annotate("text", x = 8.5, y = 27.8, 
           label = "High Point (12)", size = 1.5, color = "black") +
  annotate("text", x = 8.5, y = 25.8, 
           label = "Arkansas (4)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 8.5, y = 23.8, 
           label = "Texas (11)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 8.5, y = 21.8, 
           label = "Gonzaga (3)", size = 1.5, color = "black") +
  annotate("text", x = 8.5, y = 19.8, 
           label = "Miami FL (7)", size = 1.5, color = "black") +
  annotate("text", x = 8.5, y = 17.8, 
           label = "Purdue (2)", size = 1.5, 
           fontface = "bold", color = "black") +
  
  # Midwest Region
  annotate("label", x = 8.5, y = 15.8, 
           label = "Michigan (1)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  annotate("text", x = 8.5, y = 13.8, 
           label = "Saint Louis (9)", size = 1.5, color = "black") +
  annotate("text", x = 8.5, y = 11.8, 
           label = "Texas Tech (5)", size = 1.5, color = "black") +
  annotate("text", x = 8.5, y = 9.8, 
           label = "Alabama (4)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 8.5, y = 7.8, 
           label = "Tennessee (6)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 8.5, y = 5.8, 
           label = "Virginia (3)", size = 1.5, color = "black") +
  annotate("text", x = 8.5, y = 3.8, 
           label = "Kentucky (7)", size = 1.5, color = "black") +
  annotate("text", x = 8.5, y = 1.8, 
           label = "Iowa St. (2)", size = 1.5, 
           fontface = "bold", color = "black")

bracket_second_round
  
```

# Bracket Sweet 16
```{r}
bracket_sweet_16 <- bracket_second_round +
  # East Region
  annotate("text", x = 2.5, y = 30.8, 
           label = "Duke (1)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 2.5, y = 26.8, 
           label = "St. John's (5)", size = 1.5, color = "black") +
  annotate("text", x = 2.5, y = 22.8, 
           label = "Michigan St. (3)", size = 1.5, color = "black") +
  annotate("label", x = 2.5, y = 18.8, 
           label = "UConn (2)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  
  # South Region
  annotate("text", x = 2.5, y = 14.8, 
           label = "Iowa (9)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 2.5, y = 10.8, 
           label = "Nebraska (4)", size = 1.5, color = "black") +
  annotate("text", x = 2.5, y = 6.8, 
           label = "Illinois (3)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 2.5, y = 2.8, 
           label = "Houston (2)", size = 1.5, color = "black") +
  
  # West Region
  annotate("text", x = 7.5, y = 30.8, 
           label = "Arizona (1)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 7.5, y = 26.8, 
           label = "Arkansas (4)", size = 1.5, color = "black") +
  annotate("text", x = 7.5, y = 22.8, 
           label = "Texas (11)", size = 1.5, color = "black") +
  annotate("text", x = 7.5, y = 18.8, 
           label = "Purdue (2)", size = 1.5, 
           fontface = "bold", color = "black") +
  
  # Midwest Region
  annotate("label", x = 7.5, y = 14.8, 
           label = "Michigan (1)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  annotate("text", x = 7.5, y = 10.8, 
           label = "Alabama (4)", size = 1.5, color = "black") +
  annotate("text", x = 7.5, y = 6.8, 
           label = "Tennessee (6)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 7.5, y = 2.8, 
           label = "Iowa St. (2)", size = 1.5, color = "black")
  
bracket_sweet_16
```

# Bracket Elite 8
```{r}
bracket_elite_8 <- bracket_sweet_16 +
  # East Region
  annotate("text", x = 3.5, y = 28.8, 
           label = "Duke (1)", size = 1.5, color = "black") +
  annotate("label", x = 3.5, y = 20.8, 
           label = "UConn (2)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  
  # South Region
  annotate("text", x = 3.5, y = 12.8, 
           label = "Iowa (9)", size = 1.5, color = "black") +
  annotate("text", x = 3.5, y = 4.8, 
           label = "Illinois (3)", size = 1.5, 
           fontface = "bold", color = "black") +
  
  # West Region
  annotate("text", x = 6.5, y = 28.8, 
           label = "Arizona (1)", size = 1.5, 
           fontface = "bold", color = "black") +
  annotate("text", x = 6.5, y = 20.8, 
           label = "Purdue (2)", size = 1.5, color = "black") +
  
   # Midwest Region
  annotate("label", x = 6.5, y = 12.8, 
           label = "Michigan (1)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  annotate("text", x = 6.5, y = 4.8, 
           label = "Tennessee (6)", size = 1.5, color = "black")

bracket_elite_8
```

# Bracket Final Four
```{r}
bracket_final_4 <- bracket_elite_8 +
  annotate("label", x = 4.5, y = 24.8, 
           label = "UConn (2)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF") +
  annotate("text", x = 4.5, y = 8.8, 
           label = "Illinois (3)", size = 1.5, color = "black") +
  annotate("text", x = 5.5, y = 24.8, 
           label = "Arizona (1)", size = 1.5, color = "black") +
  annotate("label", x = 5.5, y = 8.8, 
           label = "Michigan (1)", size = 1.5, 
           fontface = "bold", color = "black", fill = "#5CCEFF")

bracket_final_4
```

# Create championship text & lines
```{r}

bracket_championship <- bracket_final_4 +
  # Centered Title at the top-middle of the canvas
  annotate("text", x = 5.0, y = 18, label = "Championship", 
           size = 4, fontface = "bold", color = "black") +
  
  # Left finalist line
  geom_segment(aes(x = 3.5, y = 15.5, xend = 4.8, yend = 15.5), color = "black", linewidth = 0.4) +
  # Right finalist line
  geom_segment(aes(x = 5.2, y = 15.5, xend = 6.5, yend = 15.5), color = "black", linewidth = 0.4) +
  
  # Place the final team names floating directly above those new lines
  annotate("label", x = 4.15, y = 16, label = "UConn (2)", size = 1.8, color = "black", fill = "#5CCEFF") +
  annotate("label", x = 5.85, y = 16, label = "Michigan (1)", size = 1.8, fontface = "bold", color = "black", fill = "#5CCEFF")
 
bracket_championship
```

# Testing Plotly Stuff
```{r}
library(plotly)

# Duke vs UConn, game ends 72-73.

duke_uconn_scores_raw <- mbb_pbp |>
  filter(game_id == "401856577") |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining) |>
  select(time_elapsed, home_score, away_score) |>
  arrange(time_elapsed) |>
  group_by(time_elapsed) |>
  slice_tail(n = 1) |>
  ungroup()

duke_uconn_db_scores <- duke_uconn_db |>
  left_join(duke_uconn_scores_raw, by = "time_elapsed") |>
  fill(home_score, away_score, .direction = "down") |>
  mutate(
    Score = paste0(
      "<span style='color:#00539B'>Duke: ", home_score, "</span> | ",
      "<span style='color:#E4002B'>UConn: ", away_score, "</span>"
    )
  )

# Plot
plot_duke_uconn <- ggplot(duke_uconn_db_scores, aes(x = time_elapsed, y = win_prob,
                                                    label = Score)) +
  geom_line(linewidth = 0.4) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_x_continuous(breaks = seq(0, 2400, by = 300)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Duke vs UConn Dynamic Bayes Win Probability",
       x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability (Duke)") +
  theme_minimal(base_size = 14)

ggplotly(plot_duke_uconn, tooltip = "label") |>
  layout(
    hoverlabel = list(
      bgcolor = "white",
      bordercolor = "black",
      font = list(color = "black")
    )
  )



# Michigan vs UConn, game ends 69-63
mich_uconn_scores_raw <- mbb_pbp |>
  filter(game_id == "401856600") |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining) |>
  select(time_elapsed, home_score, away_score) |>
  arrange(time_elapsed) |>
  group_by(time_elapsed) |>
  slice_tail(n = 1) |>
  ungroup()

mich_uconn_db_scores <- mich_uconn_db |>
  left_join(mich_uconn_scores_raw, by = "time_elapsed") |>
  fill(home_score, away_score, .direction = "down") |>
  mutate(
    Score = paste0(
      "Michigan: ", home_score, " | UConn: ", away_score, "<br>"
    )
  )

# Plot
plot_mich_uconn <- ggplot(mich_uconn_db_scores, aes(x = time_elapsed, y = win_prob,
                                                    label = Score)) +
  geom_line(linewidth = 0.4) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_x_continuous(breaks = seq(0, 2400, by = 300)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Michigan vs UConn Dynamic Bayes Win Probability",
       x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability (Michigan)") +
  theme_minimal(base_size = 14)

ggplotly(plot_mich_uconn, tooltip = "label")

```




