---
title: "data_exploration"
format: html
---

```{r}
library(tidyverse)
```

# Initial Data Collection

```{r}
# You can install using the pak package using the following code:
if (!requireNamespace('pak', quietly = TRUE)){
  install.packages('pak')
}
pak::pak("sportsdataverse/hoopR")
```

```{r}
# pbp for 2026 season
tictoc::tic()
progressr::with_progress({
  mbb_pbp <-  hoopR::load_mbb_pbp()
})
tictoc::toc()

# for all seasons: load_mbb_pbp(seasons = 2002:2026), but file is too large

print(object.size(mbb_pbp_2026), units = "auto")
```


# Initial Data Exploration

```{r}
# only keep important cols and add time_elapsed and lead_diff 
mbb_pbp_2026 <- mbb_pbp |> select(
  game_id, home_score, away_score, end_game_seconds_remaining, scoring_play, score_value, home_team_name, away_team_name, season_type, period_number, game_date 
  ) |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)
```

```{r}
# duke vs uconn game
duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_abbrev == "DUKE" & away_team_abbrev == "CONN")

```

```{r}
# There are 6275 different games from the 2026 season and there are some games that involve one non D1 school
mbb_pbp_2026 |> group_by(game_id)

```

```{r}
# home win rate
mbb_pbp_2026 |> 
  group_by(game_id) |>
  filter(time_elapsed == max(time_elapsed, na.rm = TRUE)) |>
  slice(1) |> 
  ungroup() |>
  summarise(home_win_rate = mean(home_score > away_score))

# for 2026: home win rate is 63.7% which is lower than expected (google says 68.7%)
# for 2025: home win rate is 63.4% which is lower than expected
# but not sure this calculation is correct because the home_win variable mean is 0.666 but that is by play rather than game.
```

```{r}
# duke vs uconn plot

duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_name == "Duke" & away_team_name == "UConn")

duke_uconn |>
  ggplot(aes(x = time_elapsed, y = lead_diff)) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal()
```

```{r}
# find games with least/most plays to see if there is scraping errors
mbb_pbp_2026 |>
  group_by(game_id) |>
  summarise(n_plays = n()) |>
  arrange(desc(n_plays))

# Game with least plays: the issue is that espn seems to be missing plays and not an issue with the scraping
mbb_pbp_2026 |> filter(game_id == 401824072) |> relocate(home_team_name, away_team_name)

# Game with most plays: the issue is that espn play by play glitched and it doubled up every play
mbb_pbp_2026 |> filter(game_id == 401817534) |> relocate(home_team_name, away_team_name)

```

```{r}
# what happens for an overtime game
ot_game <- mbb_pbp_2026 |> 
  filter(home_team_name == "Kentucky" & away_team_name == "Santa Clara") |>
  relocate(time_elapsed) |> arrange(desc(time_elapsed))
# for time elapsed there is one NA value
# then when in ot, end_game_seconds_remaining resets to 300 (5 mins), 
# for example then there are two very different game scores when time elapsed is 2400, it is 73-73 and 89-84

ot_game |>
  ggplot(aes(x = time_elapsed, y = lead_diff)) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal()
```


# Frequentist Logistic Regression Model

```{r}
# Get data for 2025 season
tictoc::tic()
progressr::with_progress({
  mbb_pbp_2025 <-  hoopR::load_mbb_pbp(seasons = 2025)
})
tictoc::toc()
```

```{r}
# create new vars: time_elapsed & lead_diff
mbb_pbp_2025 <- mbb_pbp_2025 |> select(
  game_id, home_score, away_score, end_game_seconds_remaining, scoring_play, score_value, home_team_name, away_team_name, season_type, period_number, game_date 
  ) |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

# create new var: home_win
mbb_pbp_2025 <- mbb_pbp_2025 |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
mbb_pbp_2025 <- mbb_pbp_2025 |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)
```

```{r}
library(broom)
library(ggeffects)
```

```{r}
# logistic regression model
log_mod <- glm(home_win ~ time_elapsed + lead_diff + time_elapsed:lead_diff,
               data = mbb_pbp_2025, family = "binomial")

log_mod |> tidy()
```

```{r}
# model predictions
preds_log_mod <- ggpredict(log_mod, terms = c("time_elapsed [0:2400 by=10]", 
                             "lead_diff [-15, -10, -5, 0, 5, 10, 15]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

# visualization
ggplot(data = preds_log_mod, aes(x = time_elapsed, y = predicted)) +
  geom_line(aes(color = factor(lead_diff)), linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff") +
  theme_minimal()
```
Plot Takeaways
- When lead_diff = 0, it starts at 67% and slowly decreases over time, so home court advantage slightly fades at time goes on
- The lines spread out over time because of the interaction term, for example a 5 point lead means more in the last minutes of the game then the first minutes.
- Home court advantage shifts all of the lines up, the positive lead diff lines are higher up than the negative lead diff lines are lower down.


```{r}
# heatmap visualization

preds_heatmap <- ggpredict(log_mod, terms = c("time_elapsed [0:2400 by=10]", 
                                               "lead_diff [-25:25 by=1]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

ggplot(preds_heatmap, aes(x = time_elapsed, y = lead_diff, fill = predicted)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5) +
  labs(x = "Time Elapsed (seconds)", 
       y = "Home Team Lead Diff",
       fill = "Home Win Prob") +
  theme_minimal()
```

```{r}
# duke vs uconn visualization

duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_name == "Duke" & away_team_name == "UConn") |>
  mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
) |>
  mutate(
    predicted = predict(log_mod, newdata = pick(everything()), type = "response"
                        ))

ggplot(duke_uconn, aes(x = time_elapsed, y = predicted)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       title = "Duke vs UConn") +
  theme_minimal()

```
The peak home team win prob was  98% right before halftime which is very similar to ESPN win probability model.


