---
title: "prob_plots"
format: html
---

# 6/26 Task: Pick a game and create a predicted probability graph through time with either MLE or the logistic model or both (dynamic bayes later).

```{r}
# Load in 2025 season pbp data
mbb_pbp <-  hoopR::load_mbb_pbp(seasons = 2025)
```

## Clean Data

```{r}
# Clean data for MLE

# create new vars: time_elapsed and lead_diff
mbb_pbp_2025 <- mbb_pbp |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

# create new var: home_win
mbb_pbp_2025 <- mbb_pbp_2025 |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
mbb_pbp_2025 <- mbb_pbp_2025 |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)

# only need these columns
mbb_pbp_2025 <- mbb_pbp_2025 |>
  select(game_id, time_elapsed, lead_diff, home_win)

# works after 1 min, 14 million rows
mbb_pbp_2025_clean <- mbb_pbp_2025 |> 
  group_by(game_id) |> 
  complete(time_elapsed = 0:2400) |> 
  fill(lead_diff, home_win, .direction = "down") |> 
  ungroup() |> 
  arrange(game_id, time_elapsed, desc(row_number())) |>
  distinct(game_id, time_elapsed, .keep_all = TRUE)


```

```{r}
# MLE calculation with cleaned data
mle_clean <- mbb_pbp_2025_clean |>
  group_by(time_elapsed, lead_diff) |>
  summarise(
    n_games = n(),
    n_wins = sum(home_win),
    win_prob = n_wins / n_games
  ) |>
  filter(n_games >= 1)
```

```{r}
# Load in 2026 season pbp data
mbb_pbp_2026 <-  hoopR::load_mbb_pbp()
```

```{r}
# Clean data for duke uconn game

# create new vars: time_elapsed and lead_diff
duke_uconn <- mbb_pbp_2026 |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score) |>
  filter(home_team_name == "Duke" & away_team_name == "UConn")

# create new var: home_win
duke_uconn <- duke_uconn |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
duke_uconn <- duke_uconn |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)

# only need these columns
duke_uconn <- duke_uconn |>
  select(game_id, time_elapsed, lead_diff, home_win)

# works after 1 min, 14 million rows
duke_uconn_clean <- duke_uconn |> 
  group_by(game_id) |> 
  complete(time_elapsed = 0:2400) |> 
  fill(lead_diff, home_win, .direction = "down") |> 
  ungroup() |> 
  arrange(game_id, time_elapsed, desc(row_number())) |>
  distinct(game_id, time_elapsed, .keep_all = TRUE)

```

## MLE Duke vs UConn

```{r}
# Join MLE probabilities to Duke vs UConn
duke_uconn_mle <- duke_uconn_clean |>
  left_join(mle_clean, by = c("time_elapsed", "lead_diff"))

# plot
ggplot(duke_uconn_mle, aes(x = time_elapsed, y = win_prob)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Duke vs UConn MLE Win Probability",
       x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability (Duke)") +
  theme_minimal()
```


## Dynamic Bayes Duke vs UConn

```{r}
# Join MLE probabilities to Duke vs UConn
duke_uconn_db <- duke_uconn_clean |>
  left_join(db_full, by = c("time_elapsed", "lead_diff"))

# plot
ggplot(duke_uconn_db, aes(x = time_elapsed, y = win_prob)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Duke vs UConn Dynamic Bayes Win Probability",
       x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability (Duke)") +
  theme_minimal()

#write_csv(duke_uconn_db, "duke_uconn_db.csv")
```


## Frequentist Logistic Regression Duke vs UConn

```{r}
# logistic regression model
log_mod <- glm(home_win ~ time_elapsed + lead_diff + time_elapsed:lead_diff,
               data = mbb_pbp_2025_clean, family = "binomial")

log_mod |> tidy()
```

```{r}
duke_uconn_log <- duke_uconn_clean |>
  mutate(
    win_prob = predict(log_mod, newdata = pick(everything()), type = "response"
                        ))

# plot
ggplot(duke_uconn_log, aes(x = time_elapsed, y = win_prob)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Duke vs UConn Logistic Win Probability",
       x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability (Duke)") +
  theme_minimal()
```


# Bayesian Logistic Regression with Ratings Duke vs. UConn

```{r}
# not working

log_full <- readRDS("bayesian_logistic_full.rds")

log_full <- log_full |>
  mutate(lead_diff = as.numeric(as.character(lead_diff)))

duke_uconn_brm <- duke_uconn_clean |>
  left_join(log_full, by = c("time_elapsed", "lead_diff"))


ggplot(duke_uconn_brm, aes(x = time_elapsed, y = predicted)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Duke vs UConn Logistic Win Probability",
       x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability (Duke)") +
  theme_minimal()
```


# 2025 and 2026 Full Datasets
```{r}
# Load in 2026 season pbp data
mbb_pbp <-  hoopR::load_mbb_pbp(seasons = 2025)

# Clean the 2025 season data
mbb_pbp_2025 <- mbb_pbp |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

# create new var: home_win
mbb_pbp_2025 <- mbb_pbp_2025 |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
mbb_pbp_2025 <- mbb_pbp_2025 |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)

# only need these columns
mbb_pbp_2025 <- mbb_pbp_2025 |>
  select(game_id, time_elapsed, lead_diff, home_win)

# works after 1 min, 14 million rows
mbb_pbp_2025_clean <- mbb_pbp_2025 |> 
  group_by(game_id) |> 
  complete(time_elapsed = 0:2400) |> 
  fill(lead_diff, home_win, .direction = "down") |> 
  ungroup() |> 
  arrange(game_id, time_elapsed, desc(row_number())) |>
  distinct(game_id, time_elapsed, .keep_all = TRUE)

write_csv(mbb_pbp_2025_clean, "mbb_pbp_2025_clean.csv")
```

```{r}
# Load in 2026 season pbp data
mbb_pbp_2026 <-  hoopR::load_mbb_pbp()

# create new vars: time_elapsed and lead_diff
mbb_pbp_2026 <- mbb_pbp_2026 |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

# create new var: home_win
mbb_pbp_2026 <- mbb_pbp_2026 |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
mbb_pbp_2026 <- mbb_pbp_2026 |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)

# only need these columns
mbb_pbp_2026 <- mbb_pbp_2026 |>
  select(game_id, time_elapsed, lead_diff, home_win)

# works after 1 min, 14 million rows
mbb_pbp_2026_clean <- mbb_pbp_2026 |> 
  group_by(game_id) |> 
  complete(time_elapsed = 0:2400) |> 
  fill(lead_diff, home_win, .direction = "down") |> 
  ungroup() |> 
  arrange(game_id, time_elapsed, desc(row_number())) |>
  distinct(game_id, time_elapsed, .keep_all = TRUE)

write_csv(mbb_pbp_2026_clean, "mbb_pbp_2026_clean.csv")
```


# Function for Predicted Probability Generation
```{r}
get_win_prob <- function(game_data, model_type, mle_clean, db_full, log_mod) {
  
  if (model_type == "mle") {
    
    result <- game_data |>
      left_join(mle_clean, by = c("time_elapsed", "lead_diff"))
    
  } else if (model_type == "db") {
    
    result <- game_data |>
      left_join(db_full, by = c("time_elapsed", "lead_diff"))
    
  } else if (model_type == "log") {
    
    result <- game_data |>
      mutate(win_prob = predict(log_mod, newdata = pick(everything()), type = "response"))
    
  } else {
    stop("model_type must be one of: 'mle', 'db', 'log'")
  }
  
  return(result)
}
```

```{r}
# Duke vs UConn Game
duke_uconn_game <- mbb_pbp_2026_clean |> filter(game_id == "401856577")

duke_uconn_mle <- get_win_prob(duke_uconn_game, "mle", mle_clean, db_full, log_mod)
duke_uconn_db  <- get_win_prob(duke_uconn_game, "db",  mle_clean, db_full, log_mod)
duke_uconn_log <- get_win_prob(duke_uconn_game, "log", mle_clean, db_full, log_mod)


# National Championship Game (Michigan vs UConn)
mich_uconn_game <- mbb_pbp_2026_clean |> filter(game_id == "401856600")

mich_uconn_mle <- get_win_prob(mich_uconn_game, "mle", mle_clean, db_full, log_mod)
mich_uconn_db  <- get_win_prob(mich_uconn_game, "db",  mle_clean, db_full, log_mod)
mich_uconn_log <- get_win_prob(mich_uconn_game, "log", mle_clean, db_full, log_mod)


# Duke vs Virginia ACC Conference Championship
duke_va_game <- mbb_pbp_2026_clean |> filter(game_id == 401851183)

duke_va_mle <- get_win_prob(duke_va_game, "mle", mle_clean, db_full, log_mod)
duke_va_db  <- get_win_prob(duke_va_game, "db",  mle_clean, db_full, log_mod)
duke_va_log <- get_win_prob(duke_va_game, "log", mle_clean, db_full, log_mod)

# Plot
# change dataframe
ggplot(duke_va_log, aes(x = time_elapsed, y = win_prob)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability") +
  theme_minimal()
```


