---
title: "Visualizing In-Game Win Probability in College Basketball Through Data Storytelling "
author: "Cooper Olney"
date: "2026-07-24"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)

mbb_pbp_2025 <-  hoopR::load_mbb_pbp(seasons = 2025)
mbb_pbp <-  hoopR::load_mbb_pbp()
```

# Summary Statistics
```{r}
mbb_pbp_2025 |> summarise(
  total_games = n_distinct(game_id),
  total_playoff_games = n_distinct(game_id[season_type == 3]),
  total_home_teams = n_distinct(home_team_id),
  total_away_teams = n_distinct(away_team_id)
  )
```

The dataset for the 2024-25 NCAA Men's Basketball season consisted of 6136 games, with 123 of those games being in the postseason. The postseason consists of many tournaments such as  March Madness, conference tournaments, and other smaller postseason tournaments such as the NIT and the College Basketball Crown. 
Meaning that during the 2024-25 season there were 6013 regular season games, in the dataset the majority of these games consisted of two NCAA Division 1 teams, but not all of the games. Taking into account that there were 362 Division 1 college basketball during this season, on average each team played about 33 regular season games.
In the dataset there were 369 different home teams, so there does consist of non-division 1 as the home team in the dataset. There is also 681 different away teams, so there are many more games where a division 1 team played at home against a non-division 1 team. (I doubt there are any games where there was D1 team) I would assume games with non-division 1 teams included are typically early in the season, as division 1 schools use these teams as warm-up games.



# Frequentist Logistic Regression Model (2025 data)
```{r}
# create new vars: time_elapsed & lead_diff
mbb_pbp_2025 <- mbb_pbp_2025 |> select(
  game_id, home_score, away_score, end_game_seconds_remaining, scoring_play, score_value, home_team_name, away_team_name, season_type, period_number, game_date 
  ) |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

# create new var: home_win
mbb_pbp_2025 <- mbb_pbp_2025 |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
mbb_pbp_2025 <- mbb_pbp_2025 |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)


library(broom)
library(ggeffects)


# logistic regression model
log_mod <- glm(home_win ~ time_elapsed + lead_diff + time_elapsed:lead_diff,
               data = mbb_pbp_2025, family = "binomial")

log_mod |> tidy()

# model predictions
preds_log_mod <- ggpredict(log_mod, terms = c("time_elapsed [0:2400 by=10]", 
                             "lead_diff [-15, -10, -5, 0, 5, 10, 15]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

# Regression visualization
ggplot(data = preds_log_mod, aes(x = time_elapsed, y = predicted)) +
  geom_line(aes(color = factor(lead_diff)), linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff") +
  theme_minimal()

# Heatmap visualization
preds_heatmap <- ggpredict(log_mod, terms = c("time_elapsed [0:2400 by=10]", 
                                               "lead_diff [-25:25 by=1]")) |>
  as.data.frame(terms_to_colnames = TRUE) |>
  as_tibble()

ggplot(preds_heatmap, aes(x = time_elapsed, y = lead_diff, fill = predicted)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5) +
  labs(x = "Time Elapsed (seconds)", 
       y = "Home Team Lead Diff",
       fill = "Home Win Prob") +
  theme_minimal()
```

*Regression Plot*
This shows how the predicted win probability of the home team changes throughout a game given seven different lead differences. With the a lead diff of 0, the win probability starts at about .67 which shows the home court advantage of college basketball. This is the only regression line that is linear because the interaction term will be not show up when lead_diff = 0. The other regression lines shows how the lead differential is weighted by time. The lead being 10 is more impactful to predicting win probability when there is one minute left compared to when the game is still in the first half. The model doesn't show how win probability should be certain, either 100% or 0% when there is no time left in the game.

*Heatmap Plot*
Now instead of only a select few lead differentials, the heatmap shows all of the combinations of lead differentials with time. 


# Duke vs Uconn
```{r}
# lead changes over time visualization
mbb_pbp_2026 <- mbb_pbp |> select(
  game_id, home_score, away_score, end_game_seconds_remaining, scoring_play, score_value, home_team_name, away_team_name, season_type, period_number, game_date 
  ) |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_name == "Duke" & away_team_name == "UConn")

duke_uconn |>
  ggplot(aes(x = time_elapsed, y = lead_diff)) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Duke Lead Diff") +
  theme_minimal()

# logistic regression with duke vs uconn
duke_uconn <- mbb_pbp_2026 |> 
  filter(home_team_name == "Duke" & away_team_name == "UConn") |>
  mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
) |>
  mutate(
    predicted = predict(log_mod, newdata = pick(everything()), type = "response"
                        ))

ggplot(duke_uconn, aes(x = time_elapsed, y = predicted)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       title = "Duke vs UConn: Duke Win Prob") +
  theme_minimal()
```

*Lead Changes Over Time Plot*
In the 2025 March Madness tournament, Duke played UConn in the Elite Eight. This plot examines how Duke's lead changed over time during this game. Just before the first half ended, Duke achieved it's largest lead of the game being 19 points. During the rest of the game UConn was able to comeback and win the game on a buzzer beating shot.

*Logistic Regression Plot*
Now applying the frequentist logistic regression model based on the 2025 data to this Duke vs. UConn game. Since Duke was considered the home team (even though the game was played at a neutral site) they started with a 67% win probability. And we can see that the win probability roughly matches the shape of the plot for the lead changes throughout the game. Right before the first half ends, the win probability peaks at around 98%, and then decreases as UConn shrinks the lead. The win probability finally goes below 50% once UConn takes the lead with zero seconds left in the game.


# MLE Heatmap
```{r}
## Clean data
# create new vars: time_elapsed and lead_diff
mbb_pbp_2026 <- mbb_pbp |>
  mutate(time_elapsed = 2400 - end_game_seconds_remaining,
         lead_diff = home_score - away_score)

# create new var: home_win
mbb_pbp_2026 <- mbb_pbp_2026 |> group_by(game_id) |>
  mutate(
    home_win = if_else(
      is.na(end_game_seconds_remaining) & home_score > away_score, 1, 0
    )) |>
  mutate(home_win = max(home_win)) |>
  ungroup(game_id)

# fill in the NAs
mbb_pbp_2026 <- mbb_pbp_2026 |> mutate(
  time_elapsed = if_else(is.na(time_elapsed), 2400, time_elapsed),
  end_game_seconds_remaining = if_else(is.na(end_game_seconds_remaining), 0, end_game_seconds_remaining)
)

# only need these columns
mbb_pbp_2026 <- mbb_pbp_2026 |>
  select(game_id, time_elapsed, lead_diff, home_win)


## MLE Calculation
mle_heatmap <- mbb_pbp_2026 |>
  group_by(time_elapsed, lead_diff) |>
  summarise(
    n_games = n(),
    n_wins = sum(home_win),
    win_prob = n_wins / n_games
  ) |>
  filter(n_games >= 1) # remove the rare situations


## MLE Heatmap
ggplot(mle_heatmap, aes(x = time_elapsed, y = lead_diff, fill = win_prob)) +
  geom_tile() +
  scale_fill_gradientn(colors = c("blue", "green", "white", "yellow", "red")) +
  labs(x = "Time elapsed", 
       y = "Score diff",
       fill = "p") +
  theme_minimal()


## Clean Data
# clean 2026 season data, 1 row for 1 second, no duplicates
mbb_pbp_2026_clean <- mbb_pbp_2026 |> 
  group_by(game_id) |> 
  complete(time_elapsed = 0:2400) |> 
  fill(lead_diff, home_win, .direction = "down") |> 
  ungroup() |> 
  arrange(game_id, time_elapsed, desc(row_number())) |>
  distinct(game_id, time_elapsed, .keep_all = TRUE)


## MLE calculation with cleaned data
mle_heatmap_clean <- mbb_pbp_2026_clean |>
  group_by(time_elapsed, lead_diff) |>
  summarise(
    n_games = n(),
    n_wins = sum(home_win),
    win_prob = n_wins / n_games
  ) |>
  filter(n_games >= 1) # remove the rare situations


## MLE Heatmap with cleaned data
ggplot(mle_heatmap_clean, aes(x = time_elapsed, y = lead_diff, fill = win_prob)) +
  geom_tile() +
  scale_fill_gradientn(colors = c("blue", "green", "white", "yellow", "red")) +
  labs(x = "Time elapsed", 
       y = "Score diff",
       fill = "p") +
  theme_minimal()
```

*Heatmap Plot*
This heatmap uses data from all games from the 2026 season, for this data each row is an event and there can be multiple events a second. Overall the plot follows the general trends where a large lead results in high win prob and a large deficit results in a low win prob. The major issue with the MLE model is that for rare situations, for example, games with a 10 point deficit 2 minutes into the game doesn't happen often so there might be inaccurate win probability. Which is why there are some red tiles in spots where it doesn't make sense for the win probability to be extremely high.

*Clean Heatmap Plot*
This heatmap also uses data from the 2026 season, but each row is 1 second and the duplicate events are removed, resulting in more total rows of data. This heatmap looks very similar to the previous heatmap, since now more combinations of score and time are covered the plot looks smoother. But there is still some inaccurate win probabilities results. Around 800 seconds and a score diff of -30, there are some dark red tiles meaning that the win probability was 100%, this is because one game occurred where the home team was down 30 points 800 seconds into the game, but they ended up winning which causes these outlying red tiles.


# Dynamic Bayes Heatmap
```{r}
db_full <- readRDS("dynamic_bayes_full.rds")

ggplot(db_full, aes(x = time_elapsed, y = lead_diff, fill = win_prob)) +
  geom_tile() +
  scale_fill_gradientn(colors = c("blue", "green", "white", "yellow", "red")) +
  labs(
    x = "Time elapsed (seconds)", 
    y = "Score differential",
    fill = "p"
  ) +
  theme_minimal()
```

*Heatmap*
This heatmap uses data from the 2025 season, and for every row is 1 second of game data. This heatmap compares very well to the heatmap used in the referenced research paper where they used data from multiple seasons, so this heatmap is able to capture the overall predicted win probability in each "window". A window is +/- 3 seconds and 2 points from the current time and score combination. This is helps with the rare situations, so now this heatmap doesn't contain very high or low win probabilities where they shouldn't be. Also the dynamic bayes heatmap uses beta priors, so the win probability won't solely rely on the data, this results in the predicted win probability to never be exactly 100% or 0%. The dynamic bayes model and heatmap has similar calculations to the MLE method, but it is an overall improvement.


# Probability Plots
```{r}
# Dynamic Bayes: Duke vs UConn
ggplot(duke_uconn_db, aes(x = time_elapsed, y = win_prob)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability") +
  theme_minimal()

# Dynamic Bayes: Michigan vs UConn
ggplot(mich_uconn_db, aes(x = time_elapsed, y = win_prob)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability") +
  theme_minimal()

# Dynamic Bayes: Duke vs Virginia
ggplot(duke_va_db, aes(x = time_elapsed, y = win_prob)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  geom_vline(xintercept = 1200, linetype = "dashed", color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Time Elapsed (seconds)",
       y = "Predicted Home Win Probability") +
  theme_minimal()
```

*Predicted Probability Plots*
These three plots will be in the closeread document, using a function you can easily create one of these plots for any game in the dataset and using one of the these three methods: MLE, Dynamic Bayes, or Logistic Regression. These predicted probability plots map the home team's predicted win probability throughout the entire game. The plots listed all use the Dynamic Bayes method to predicted win probability, and they cover three interesting games from the 2026 college basketball season. First, Duke vs. UConn in the Elite Eight saw Duke reach a 19 point lead and reaching a win probability of 98% in the first half. But a large comeback and a buzzer beating 3-pointer saw UConn win the game and move on in the tournament. Second, Michigan vs. UConn in the National Championship game, this game isn't too intersting, but it shows how Michigan retained a small lead throughout the game and how their win probability increased as they got further into the game. Lastly, Duke vs. Virginia in the ACC Conference Championship, this was a very back and forth game, where neither team could keep their lead. Despite being down multiple times in this game, Duke had over a 50% win probability most of the time, showing how the model incorporates home court advantage in predicting home team win probability.

# Bayesian Logistic Regression w/ Ratings
```{r}
brm_full <- readRDS("brm_full.rds")

min_lead_diff <- min(subset_games$lead_diff)
max_lead_diff <- max(subset_games$lead_diff)

min_rating_diff <- min(subset_games$rating_diff)
max_rating_diff <- max(subset_games$rating_diff)

# Build prediction grid in original units, then scale to match the model
pred_grid <- expand.grid(
  time_elapsed = seq(0, 2400, by = 10),
  lead_diff    = c(-15, -10, -5, 0, 5, 10, 15),
  rating_diff  = 0   # hold rating_diff at 0 (evenly matched teams)
) |>
  mutate(
    time_scaled   = (time_elapsed - 0) / (2400 - 0),
    lead_scaled   = (lead_diff - min_lead_diff) / (max_lead_diff - min_lead_diff),
    rating_scaled = (rating_diff - min_rating_diff) / (max_rating_diff - min_rating_diff)
  )

# Get posterior predicted probabilities
preds_brm_mod_rank_scaled <- pred_grid |>
  mutate(
    predicted = posterior_epred(brm_mod_rank, newdata = pred_grid) |>
      colMeans(),
    # optional: get credible intervals too
    lower = posterior_epred(brm_mod_rank, newdata = pred_grid) |>
      apply(2, quantile, 0.025),
    upper = posterior_epred(brm_mod_rank, newdata = pred_grid) |>
      apply(2, quantile, 0.975)
  )

# plot
ggplot(preds_brm_mod_rank_scaled, aes(x = time_elapsed, y = predicted, 
                          color = factor(lead_diff))) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = factor(lead_diff)),
              alpha = 0.25, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_colour_viridis_d(option = "D") +
  scale_fill_viridis_d(option = "D") +
  labs(x = "Time Elapsed (seconds)", 
       y = "Predicted Home Win Probability",
       color = "Lead Diff", 
       fill = "Lead Diff") +
  theme_minimal()
```

*Regression Plot*
This plot shows how the predicted win probability of the home team changes throughout a game given seven different lead differences. In the first 300 seconds of the game, the order or the lead differences regression lines is flipped to how you would expect them to be. This is likely due to extrapolation, where there is very limited data on large lead differences in the first few minutes of the game, so the model struggles to predict win probabilities during that time. Looking at the lead diff of 0, when time elapsed is 0, the win probability is slightly below 50% which is surprising because all of the other models have the starting win probability of the home team in the mid 60s, so this model shows no home team advantage. As the game goes on, the regression lines spread out towards either 100% or 0% because there is less time for a comeback. The model could be revised to remove data from the first few minutes, so there is no extrapolation.




